Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra

### 2015-06-04 ms.tex
Emphasize usefulness for power spectrum estimators
From 2fd11fe:
============

In eor.bib,

(*) Add reference to Dillon et al.(2015)

In ms.tex,

(*) Modify Abstract and Sections 5 and 6 to be specific about
implications for antenna design
(*) Add a paragraph to emphasize optimum power spectrum estimation
techniques must take the wide-field "pitchfork" into account
(*) Insert updated .bbl file contents in bibliography

We confirm our recent prediction of the ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures from the simulation which predicted this feature were validated against Murchison Widefield Array (MWA) data but this key {\it pitchfork} signature was close to the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna designs that suppress sensitivity near the horizon. Simple models for antenna apertures proposed for future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array indicate they should suppress foreground leakage from the {\it pitchfork} by $\sim 40$~dB relative to the MWA, and significantly increase the likelihood of cosmological signal detection in three-dimensional power spectrum.

@@

We note that reducing uncertainties will only change the relative strength of the {\it pitchfork} signature in our model. However, the effects giving rise to this signature are generic to all wide-field measurements. Thus, the extremely high significance detection of foreground emission near the horizon limits is a robust confirmation of the predicted wide-field measurement phenomenon.

@@

With the SKA at low frequencies, deploying ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas is being considered\footnote{Documents at \url{https://www.skatelescope.org/key-documents/} on SKA re-baselining}$^,$\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.}. The voltage beam of the station will be obtained by a phased addition of the dipole responses. Since the power pattern of a baseline consisting of two stations will be effectively a product of two voltage beams, it will tend to homogenize the fluctuations in the power patterns of individual stations. Even with unconstrained, and hence sub-optimal, realizations of these phased array stations, we estimate the typical horizon response of these beams to be $-30$~dB to $-40$~dB compared to zenith, yielding a response that is at least 10~dB better than the MWA. Constrained optimization of antenna layouts of stations is under active study and could yield even better responses.

From these simple antenna aperture models, HERA and SKA should mitigate leakage from the {\it pitchfork} into the {\it EoR window} by $\sim 40$~dB relative to the MWA when expressed in units of delay power spectrum. This will significantly increase the likelihood of detecting the cosmological H~{\sc i} power spectrum in {\it voxels} in three-dimensional $k$-space. The precise beam responses and the resulting foreground leakage, especially from near the horizon, will prove to be critical and timely inputs to the actual aperture designs of HERA dishes and SKA stations which are nearing their final stages.

@@

Using deeper MWA data, we have confirmed with high significance the earlier prediction that wide-field EoR measurements suffer significant foreground contamination from near the horizon. This has important implications for instrument design and data analysis of future instruments such as HERA and SKA. Careful aperture designs that suppress the response near the horizon will significantly enhance the likelihood of detecting the cosmological signal in three-dimensional power spectrum. Precise modeling is thus required to gain a complete understanding of the characteristics of the cosmological signal and the foregrounds.

@@

@@

\end{thebibliography}

\end{document}

>>>>

We confirm our recent prediction of the ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures from the simulation which predicted this feature were validated against Murchison Widefield Array (MWA) data but this key {\it pitchfork} signature was close to the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna designs that suppress sensitivity near the horizon. Simple models for antenna apertures proposed for future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array indicate they should suppress foreground leakage from the {\it pitchfork} by $\sim 40$~dB relative to the MWA, and significantly increase the likelihood of cosmological signal detection in these critical Fourier modes in the three-dimensional power spectrum.

@@

We note that reducing uncertainties will only change the relative strength of the {\it pitchfork} signature in our model. However, the effects giving rise to this signature are generic to all wide-field measurements. Thus, the extremely high significance detection of foreground emission near the horizon limits is a robust confirmation of the predicted effects of wide-field measurements.

@@

With the SKA at low frequencies, deploying ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas is being considered\footnote{Documents at \url{https://www.skatelescope.org/key-documents/} on SKA re-baselining}$^,$\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.}. The voltage beam of the station will be obtained by a phased addition of the dipole responses. Since the power pattern of a baseline consisting of two stations will be effectively a product of two voltage beams, it will tend to homogenize the fluctuations in the power patterns of individual stations. Even with simple models of such phased array stations, we estimate the typical horizon response of these beams to be $-30$~dB to $-40$~dB compared to zenith, yielding a response that is at least 10~dB better than the MWA. Targeted optimization of antenna layouts of stations is under active study and could yield even better responses.

From these simple antenna aperture models, HERA and SKA should mitigate leakage from the {\it pitchfork} into the {\it EoR window} by $\sim 40$~dB relative to the MWA when expressed in units of delay power spectrum. This will significantly increase the likelihood of detecting the cosmological H~{\sc i} power spectrum in sensitive three-dimensional $k$-modes adjacent to the {\it pitchfork}. The precise beam responses and the resulting foreground leakage, especially from near the horizon, will prove to be critical and timely inputs to the actual aperture designs of HERA dishes and SKA stations which are nearing their final stages.

@@

Using deeper MWA data, we have confirmed with high significance the earlier prediction of a characteristic {\it pitchfork} morphology wherein wide-field EoR measurements suffer from significant foreground contamination from near the horizon. This has important implications for instrument design and data analysis of future instruments such as HERA and SKA.

Careful aperture designs that suppress the response near the horizon, and hence also the leakage from the {\it pitchfork}, will significantly avoid contamination in $k$-modes considered critical for cosmological signal detection. Precise modeling is thus required to gain a complete understanding of the characteristics of the cosmological signal and the foregrounds.

Foreground and instrument models serve as inputs for power spectrum estimation techniques \citep[for e.g.,][]{liu14a,liu14b,dil15}. Input models which ignore wide-field effects will provide over-valued weights in and around the {\it pitchfork} modes, and hence result in sub-optimal results. Thus, confirmation of the {\it pitchfork} effect has important implications for power spectrum estimation.

@@

\bibitem[{{Dillon} {et~al.}(2015){Dillon}, {Neben}, {Hewitt}, {Tegmark},
  {Barry}, {Beardsley}, {Bowman}, {Briggs}, {Carroll}, {de Oliveira-Costa},
  {Ewall-Wice}, {Feng}, {Greenhill}, {Hazelton}, {Hernquist}, {Hurley-Walker},
  {Jacobs}, {Kim}, {Kittiwisit}, {Lenc}, {Line}, {Loeb}, {McKinley},
  {Mitchell}, {Morales}, {Offringa}, {Paul}, {Pindor}, {Pober}, {Procopio},
  {Riding}, {Sethi}, {Udaya Shankar}, {Subrahmanyan}, {Sullivan},
  {Thyagarajan}, {Tingay}, {Trott}, {Wayth}, {Webster}, {Wyithe}, {Bernardi},
  {Cappallo}, {Deshpande}, {Johnston-Hollitt}, {Kaplan}, {Lonsdale},
  {McWhirter}, {Morgan}, {Oberoi}, {Ord}, {Prabu}, {Srivani}, {Williams}, \&
  {Williams}}]{dil15}
{Dillon}, J.~S., {Neben}, A.~R., {Hewitt}, J.~N., {et~al.} 2015, ArXiv
  e-prints, arXiv:1506.01026

@@

\end{thebibliography}
\end{document}
